---
title: "NYC Shooting Project"
author: "SSH"
date: "2023-02-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Peer graded assignment NYPD shooting incident

The objective of this assignment is to test the capability of the student in applying the data science concepts taught and provide effective outcomes.

Step 1:
Import data from the Server

```{r get_nypd_data}
library(tidyverse)
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/"
file_names <- c("rows.csv")
urls <- str_c(url_in,file_names)
```

```{r import_data,message=FALSE}
NYPD_Shooting <- read_csv(urls)
```
 
Step 2:
Tidying the data:
 Identified the suitable fields for the analysis and removed the unwanted fields. Changed the date fields in accordance with the R suited format.
 ```{r tidy_NYPD,echo=TRUE}
 ##Removing the least relevant fields
 library(tidyr)
 library(tidyverse)
 library(lubridate)
 library(dplyr)
 library(ggplot2)
 NYPD_Shooting <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
 knitr::kable(head(NYPD_Shooting))
  nypd_cleansed <- drop_na(NYPD_Shooting) %>% select(-c(INCIDENT_KEY, LOCATION_DESC, X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat, JURISDICTION_CODE, OCCUR_TIME))
  ##Changing the date to the convenience
  nypd_cleansed <- nypd_cleansed %>% mutate(OCCUR_DATE = mdy(OCCUR_DATE)) 
  ##Converting the boolean values to integers
  nypd_cleansed$STATISTICAL_MURDER_FLAG[nypd_cleansed$STATISTICAL_MURDER_FLAG == "TRUE"] <- 1
  nypd_cleansed$STATISTICAL_MURDER_FLAG[nypd_cleansed$STATISTICAL_MURDER_FLAG == "FALSE"] <- 0
  nypd_boro <- nypd_cleansed  %>% group_by(BORO, OCCUR_DATE) %>% summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>% select(BORO, OCCUR_DATE, STATISTICAL_MURDER_FLAG) %>% ungroup()
  nypd_boro$cummurder <- ave(nypd_boro$STATISTICAL_MURDER_FLAG,nypd_boro$BORO,FUN=cumsum)
  nypd_boro['shooting']=1
  nypd_boro$cumshooting <- ave(nypd_boro$shooting,nypd_boro$BORO, FUN = cumsum)
  nypd_boro$murderpercent <- with(nypd_boro, cummurder/cumshooting *100)
  #show the final data for the anlaysis
  knitr::kable(head(nypd_boro))
  ```
Step 3:
Data Analysis
Aggregated the required measures based on the suitable dimensions such as BORO,Date and Flag.
   ```{r}
   aggregate(nypd_boro$STATISTICAL_MURDER_FLAG, by=list(BORO = nypd_boro$BORO), FUN=sum)
aggregate(nypd_boro$shooting, by=list(BORO = nypd_boro$BORO), FUN=sum)
city <- "BRONX"
nypd_murder_boro_BRONX <- nypd_boro %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_BRONX))
city <- "BROOKLYN"
nypd_murder_boro_BROOKLYN <- nypd_boro %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_BROOKLYN))
city <- "STATEN ISLAND"
nypd_murder_boro_STATENISLAND <- nypd_boro %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_STATENISLAND))
city <- "MANHATTAN"
nypd_murder_boro_MANHATTAN <- nypd_boro %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_MANHATTAN))
city <- "QUEENS"
nypd_murder_boro_QUEENS <- nypd_boro %>%
     filter(BORO == city) %>%
     group_by(BORO, OCCUR_DATE) %>%
     #summarize(STATISTICAL_MURDER_FLAG = STATISTICAL_MURDER_FLAG) %>%
     select(BORO, OCCUR_DATE, shooting, cumshooting, STATISTICAL_MURDER_FLAG, cummurder, murderpercent) %>%
     ungroup()
knitr::kable(tail(nypd_murder_boro_QUEENS))
```
Step 4:
Applying Linear model on the data and Visualization
   ```{r}
mod <- lm(cumshooting ~ cummurder, data = nypd_boro)
summary(mod)
nypd_boro %>% slice_min(cumshooting)
nypd_boro %>% slice_max(cumshooting)
x_grid <- seq(0, 3000)
new_df <- tibble(cumshooting = x_grid)
nypd_pred <- nypd_boro %>% mutate(pred = predict(mod))
# nypd_pred
nypd_pred %>% ggplot() +
  geom_point(aes(x = OCCUR_DATE, y=cumshooting), color= "green")+
  geom_point(aes(x = OCCUR_DATE, y = pred), color = "red")

#Visualization of data

nypd_boro %>%
  ggplot(aes(x = OCCUR_DATE, y=cummurder, group=BORO, color=BORO))+
  geom_line()
nypd_boro %>%
  ggplot(aes(x = OCCUR_DATE, y=cumshooting, group=BORO, color=BORO))+
  geom_line()
nypd_boro %>%
  ggplot(aes(x = OCCUR_DATE, y=murderpercent, group=BORO, color=BORO))+
  geom_line()
```
Step 5:
Adding Bias Identification
Being foreigner I couldn't imagine the incident in the way it happened and that's a potential Bias.I have great fear of shooting and disbelief of the society in which it is carried out.Had to do lot of studies to understand the incident and this could lead me to the way it was portrayed.

 
 
